---
title: "Threat_Special_Issue_Study2"
author: "Gautier Lucas O.S."
date: "2025-01-22"
bibliography: [bibliography.bib, book.bib, packages.bib]
link-citations: yes
output:
  html_document: default
  word_document: default
---

```{r Package loading-Exp2, include = FALSE}

rm(list=ls()) # Clear the environment

require(pacman)

p_load(tinytex, knitr, kableExtra, readr, dplyr, tidyverse, Hmisc, mada, party, pdp, psych, lmerTest, mice, VIM, missForest, lavaan, semPlot, lattice, sjPlot, sjtable2df, broom.mixed, performance, insight, MVN, ggplot2, rsq, viridis, hrbrthemes)

load("EnvironmentSaving/Part1_Menace/Task_Locate_2025-01-10.RData")

load("EnvironmentSaving/Part1_Menace/Survey_Analysis_Location_2024-10-08.RData")

# Load Model Comparison for RT
load(file = c("EnvironmentSaving/Part1_Menace/Model_Comparison/Locate_Comp_Model_1_Item.RData"))
load(file = c("EnvironmentSaving/Part1_Menace/Model_Comparison/Locate_Comp_Model_3.RData"))
load(file = c("EnvironmentSaving/Part1_Menace/Model_Comparison/Locate_Comp_Model_Acc_1.RData"))
load(file = c("EnvironmentSaving/Part1_Menace/Model_Comparison/Locate_Comp_Model_Acc_2.RData"))

load(file = c("EnvironmentSaving/Part1_Menace/Model_Comparison/Locate_Models/RT_ModComp01_Item.RData"))
load(file = c("EnvironmentSaving/Part1_Menace/Model_Comparison/Locate_Models/RT_ModComp01.RData"))


rounding2 <- function(r){
  round(r, digits = 2) 
}

rounding0 <- function(r){
  round(r, digits = 0) 
}

pValue <- function(p){
  ifelse((p)<= 0.001 ,"< 0.001", 
         ifelse((p)<= 0.01 ,"< 0.01", 
                paste0("= ", round((p), digit = 3))))
}

```


## Method

### Population

The total sample for this study consists of *N* = `r (nlevels(factor(df_Location_Full$response_id)))` french participants, recruited via the *CrowdPanel* platform and compensated for their participation. 
The final sample size, after excluding participants reporting sound issues (*n* = `r (nlevels(N_No_Sound$response_id))`), is *N* = `r (nlevels(df_Locate$response_id))`. 
The a priori calculated sample size for this experiment was identical to that determined for Study 1. 
The sample includes `r table(df_Survey_Wide_Location$Gender_str)[["Woman"]]` women and `r table(df_Survey_Wide_Location$Gender_str)[["Man"]]` men (*Mean*~Age~ = `r rounding2(mean(df_Survey_Wide_Location$Age))`, *SD*~Age~ = `r rounding2(sd(df_Survey_Wide_Location$Age))`, *min* = `r min(df_Survey_Wide_Location$Age)`, *max* = `r max(df_Survey_Wide_Location$Age)`). 
Participants primarily report occupations as employees (*n* = `r table(df_Survey_Wide_Location$CSP_str)[["Employee"]]`), intellectual professions (*n* = `r table(df_Survey_Wide_Location$CSP_str)[["IntellectualProfession"]]`), or intermediate professions (*n* = `r table(df_Survey_Wide_Location$CSP_str)[["IntermediateProfession"]]`). 
Finally, `r table(df_Survey_Wide_Location$PTSD_Diag)[["Diag"]]` participants report a *PTSD* score exceeding the clinical diagnostic threshold (i.e., > 44).

### Material and procedure

Material and procedure for this experiment were identical to those in Study 1, except that participants performed the spatial cueing task with a simple search instruction rather than a Compound search instruction.
In this task, a colored target, either red (*n* = `r table(df_Survey_Wide_Location$Target_Color)[["red"]]`) ou green (*n* = `r table(df_Survey_Wide_Location$Target_Color)[["green"]]`), could randomly appear in one of four position markers. 
In Study 2, participants were required to indicate the location of the target's appearance rather than categorizing it.
Participants responded using the keys *Y*, *H*, *B*, and *G*, corresponding to the target's appearance at the *top*, *bottom*, *right*, or *left* of the screen, respectively.
Apart from the change in the general task instruction, material and procedure of Study 2 remained unchanged (see Study 1 for a detailed description).
Material for this study is available on OSF (<https://osf.io/jkt9m/>).


## Results

### Behavioral data

#### Statistical Approach

In the second study, model comparison procedures similar to those in Study 1 were applied to select appropriate mixed models. Identical exclusion criteria were also applied to the final models.

The final model used to predict response times is as follows:

```
Response Time ~ Congruency * Validity * Condition + 
     (Congruency * Validity + Congruency * Condition || Participant)
```

For predicting response accuracy, the final model selected is:

```
Accuracy ~ Congruency * Validity * Condition + 
      (Validity + Congruency * Condition || Participant) +
      (Validity || Target-Type)
```


#### Effect of Threat

Regarding the effect of threat on response times, our analysis reveals a significant main effect of the experimental condition (*b* = `r round(coefs_RT_ModFinal_02["Condition_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Condition_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Condition_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_02["Condition_C", "Pr...t.."])`). 
In the *Threat* condition, participants are significantly faster at locating the target (*M* =
`r round(Mean_RT_Condition["Threat", "mean"], digits = 0)` ms,
*SE* = `r round(Mean_RT_Condition["Threat", "se"], digits = 2)`) compared to the *control* condition (*M* =
`r round(Mean_RT_Condition["Control", "mean"], digits = 0)` ms,
*SE* = `r round(Mean_RT_Condition["Control", "se"], digits = 2)`). 
Contrary to our hypotheses, this analysis did not reveal a significant interaction of the experimental condition with other variables in the model. 
Specifically, there was no significant interaction with validity (*b* = `r round(coefs_RT_ModFinal_02["Validity_C:Condition_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Validity_C:Condition_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Validity_C:Condition_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_02["Validity_C:Condition_C", "Pr...t.."])`), nor with congruency (*b* = `r round(coefs_RT_ModFinal_02["Congruency_C:Condition_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Congruency_C:Condition_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Congruency_C:Condition_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_02["Congruency_C:Condition_C", "Pr...t.."])`), nor the 3-way interaction (*b* = `r round(coefs_RT_ModFinal_02["Congruency_C:Validity_C:Condition_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Congruency_C:Validity_C:Condition_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Congruency_C:Validity_C:Condition_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_02["Congruency_C:Validity_C:Condition_C", "Pr...t.."])`). 
However, an equivalence analysis does not allow us to conclude equivalence for the 3-way interaction effect with a standardized effect size *d* = `r rounding2(d_SESOI_Fixef_Posteriori)` (*SESOI* = 9.76 ms, *t*(`r rounding2(lower_RT_Locate$df)`) = `r rounding2(lower_RT_Locate$'t value') `, *p* `r pValue(lower_RT_Locate$'Pr(>|t|)'/2)`). 
Although the results of this analysis suggest that threat does not moderate attentional capture, the analyses nonetheless show that participants' response speed is improved when distress screams are heard, compared to when no screams are heard.



Regarding the accuracy of responses, our analysis does not reveal a significant main effect of the threat (*b* = `r round(coefs_Accuracy_ModFinal["Condition_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Condition_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Condition_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Condition_C", "Pr...z.."])`). 
This analysis also does not show a significant interaction effect with the validity of the cue (*b* = `r round(coefs_Accuracy_ModFinal["Validity_C:Condition_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Validity_C:Condition_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Validity_C:Condition_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Validity_C:Condition_C", "Pr...z.."])`). 
However, it does show a marginal interaction effect between threat and the congruency of the cue (*b* = `r round(coefs_Accuracy_ModFinal["Congruency_C:Condition_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Congruency_C:Condition_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Congruency_C:Condition_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Congruency_C:Condition_C", "Pr...z.."])`), as well as a marginal 3-way interaction as predicted by our hypotheses (*b* = `r round(coefs_Accuracy_ModFinal["Congruency_C:Validity_C:Condition_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Congruency_C:Validity_C:Condition_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Congruency_C:Validity_C:Condition_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Congruency_C:Validity_C:Condition_C", "Pr...z.."])`). 



In summary, these results suggest that the manipulation of threat does not have a main effect on the overall performance of participants, even though participants tend to respond more quickly in the threatening situation. 
However, screams seems to influence the effect of congruency as well as the relationship between congruency and validity on performance.
First, the *Threat* condition is associated with an increasing effect of congruency. 
In fact, the difference in accuracy between congruent and incongruent trials tends to be amplified in the *Threat* block (*b* =
`r round(coefs_Accuracy_SimpEffect_Threat["Congruency_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_SimpEffect_Threat, type = "ml1")[["Congruency_C"]], digit =0)`)
=
`r round(coefs_Accuracy_SimpEffect_Threat["Congruency_C", "z.value"], digits = 2)`,
*p* `r  pValue(coefs_Accuracy_SimpEffect_Threat["Congruency_C", "Pr...z.."])`)
compared to the *Control* block (*b* =
`r round(coefs_Accuracy_SimpEffect_Ctrl["Congruency_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_SimpEffect_Ctrl, type = "ml1")[["Congruency_C"]], digit =0)`)
=
`r round(coefs_Accuracy_SimpEffect_Ctrl["Congruency_C", "z.value"], digits = 2)`,
*p* `r  pValue(coefs_Accuracy_SimpEffect_Ctrl["Congruency_C", "Pr...z.."])`).
Additionally, the interaction between Congruency and Validity tends to be amplified in the presence of distress screams compared to a control condition. Thus, the screams lead to an increase in the contingent capture effect (*b* =
`r round(coefs_Accuracy_SimpEffect_Threat["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_SimpEffect_Threat, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`)
=
`r round(coefs_Accuracy_SimpEffect_Threat["Congruency_C:Validity_C", "z.value"], digits = 2)`,
*p* `r  pValue(coefs_Accuracy_SimpEffect_Threat["Congruency_C:Validity_C", "Pr...z.."])`) compared to a situation with no sounds (*b* =
`r round(coefs_Accuracy_SimpEffect_Ctrl["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_SimpEffect_Ctrl, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`)
=
`r round(coefs_Accuracy_SimpEffect_Ctrl["Congruency_C:Validity_C", "z.value"], digits = 2)`,
*p* `r  pValue(coefs_Accuracy_SimpEffect_Ctrl["Congruency_C:Validity_C", "Pr...z.."])`). 
Although interaction effects are marginal, these results align with the hypothesis of a reinforcement of goal-driven attentional priorities, suggesting that threat amplifies processing of congruent cues while diminishing processing of incongruent cues.



### Exploratory analyses

Similar to Study 1, we aggregated the *Threat* and *Vocalize* conditions and compared their effects with the *Control* condition. 
This decision was motivated by the absence of a significant 3-way interaction when comparing the *Threat* and *Vocalize* blocks, both for reaction times (*b* = `r round(coefs_RT_ModFinal_SoundComp_02["Congruency_C:Validity_C:Sound_Type", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_SoundComp_02, type = "ml1")[["Congruency_C:Validity_C:Sound_Type"]], digit =0)`) =
`r round(coefs_RT_ModFinal_SoundComp_02["Congruency_C:Validity_C:Sound_Type", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_SoundComp_02["Congruency_C:Validity_C:Sound_Type", "Pr...t.."])`) and for accuracy (*b* = `r round(coefs_Accuracy_ModFinal_SoundComp["Congruency_C:Validity_C:Sound_Type", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal_SoundComp, type = "ml1")[["Congruency_C:Validity_C:Sound_Type"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal_SoundComp["Congruency_C:Validity_C:Sound_Type", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal_SoundComp["Congruency_C:Validity_C:Sound_Type", "Pr...z.."])`). 
However, in this study, an equivalence analysis on reaction times failed to conclude that the impact of sounds on contingent capture was marginally smaller than a non-standardized effect size of 9.76 ms (*t*(`r rounding2(upper_SoundComp$df)`) = `r rounding2(upper_SoundComp$'t value') `, *p* `r pValue(upper_SoundComp$'Pr(>|t|)'/2)`).


#### Effect of Unpredictable Sounds 

In terms of response times, exploratory analyses reveal fixed effects that are identical to the results reported without considering the *Vocalize* block. This analysis also shows a marginal 3-way interaction effect (*b* = `r round(coefs_RT_ModFinal_ThTo_02["Congruency_C:Validity_C:Sounds_Presence_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_02, type = "ml1")[["Congruency_C:Validity_C:Sounds_Presence_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_02["Congruency_C:Validity_C:Sounds_Presence_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_ThTo_02["Congruency_C:Validity_C:Sounds_Presence_C", "Pr...t.."])`). 
The simple effects suggest that the contingent capture effect is amplified when sounds (i.e., distress screams or human vocalizations) are unpredictably heard during the trial block (*b* = `r round(coefs_RT_ModFinal_ThTo_SimpEffect_Sound["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_SimpEffect_Sound, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_SimpEffect_Sound["Congruency_C:Validity_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_ThTo_SimpEffect_Sound["Congruency_C:Validity_C", "Pr...t.."])`) compared to a block where no sound is heard (*b* = `r round(coefs_RT_ModFinal_ThTo_SimpEffect_NoSound["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_SimpEffect_NoSound, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_SimpEffect_NoSound["Congruency_C:Validity_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_ThTo_SimpEffect_NoSound["Congruency_C:Validity_C", "Pr...t.."])`). 
Thus, compared to the *Control* block, when participants hear sounds, the validity effect is increased by `r rounding2(abs(coefs_RT_ModFinal_ThTo_02["Congruency_C:Validity_C:Sounds_Presence_C", "Estimate"]))` ms in trials with a target/cue match compared to trials with a target/cue mismatch.


More specifically, in the *Control* condition, when the target and cue are congruent, the difference in response times between valid and invalid trials is `r rounding2(abs(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_NoSound["Validity_C", "Estimate"]))` ms (*b* = `r round(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_NoSound["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_SimpEffect_Cong_NoSound, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_NoSound["Validity_C", "t.value"], digits = 2)`, *p* `r pValue(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_NoSound["Validity_C", "Pr...t.."])`), while this difference is larger (i.e., `r rounding2(abs(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_Sound["Validity_C", "Estimate"]))`ms) when unpredictable screams are heard (*b* = `r round(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_Sound["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_SimpEffect_Cong_Sound, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_Sound["Validity_C", "t.value"], digits = 2)`, *p* `r pValue(coefs_RT_ModFinal_ThTo_SimpEffect_Cong_Sound["Validity_C", "Pr...t.."])`). 
Regarding incongruent cues, analysis does not suggest a reduction in the validity effect due to the occurence of screams. Thus, when the target and cue are incongruent, the validity effect is not significant in the *Control* condition (*b* = `r round(coefs_RT_ModFinal_ThTo_SimpEffect_Incong_NoSound["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_SimpEffect_Incong_NoSound, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_SimpEffect_Incong_NoSound["Validity_C", "t.value"], digits = 2)`, *p* `r pValue(coefs_RT_ModFinal_ThTo_SimpEffect_Incong_NoSound["Validity_C", "Pr...t.."])`), whereas this effect is significant and of larger size in the blocks where sounds are heard (*b* = `r round(coefs_RT_ModFinal_ThTo_SimpEffect_Incong_Sound["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_ThTo_SimpEffect_Incong_Sound, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_ThTo_SimpEffect_Incong_Sound["Validity_C", "t.value"], digits = 2)`, *p* `r pValue(coefs_RT_ModFinal_ThTo_SimpEffect_Incong_Sound["Validity_C", "Pr...t.."])`). 



Regarding response accuracy, exploratory analyses replicate the main and interaction effects found without considering the *Vocalize* block. 
Additionally, the analyses reveal a significant 3-way interaction effect  (*b* = `r round(coefs_Accuracy_ModFinal_ThTo["Congruency_C:Validity_C:Sounds_Presence_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal_ThTo, type = "ml1")[["Congruency_C:Validity_C:Sounds_Presence_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal_ThTo["Congruency_C:Validity_C:Sounds_Presence_C", "z.value"], digits = 2)`, *OR* = `r round(OR_Accuracy_ModFinal_ThTo["Congruency_C:Validity_C:Sounds_Presence_C", "estimate"], digits = 2)`, IC à 95% [`r round(OR_Accuracy_ModFinal_ThTo["Congruency_C:Validity_C:Sounds_Presence_C", "conf.low"], digits = 2)`, `r round(OR_Accuracy_ModFinal_ThTo["Congruency_C:Validity_C:Sounds_Presence_C", "conf.high"], digits = 2)`], *p* `r  pValue(coefs_Accuracy_ModFinal_ThTo["Congruency_C:Validity_C:Sounds_Presence_C", "Pr...z.."])`). 
According to this result, the contingent capture effect is enhanced when unpredictable screams are heard compared to a block without sounds.
Simple effects reveal that the contingent capture effect is stronger in blocks with sounds (*b* = `r round(coefs_Acc_ModFinal_ThTo_SimpEffect_Sound["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Acc_ModFinal_ThTo_SimpEffect_Sound, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`) =
`r round(coefs_Acc_ModFinal_ThTo_SimpEffect_Sound["Congruency_C:Validity_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Acc_ModFinal_ThTo_SimpEffect_Sound["Congruency_C:Validity_C", "Pr...z.."])`) compared to a block with no sounds (*b* = `r round(coefs_Acc_ModFinal_ThTo_SimpEffect_NoSound["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Acc_ModFinal_ThTo_SimpEffect_NoSound, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`) =
`r round(coefs_Acc_ModFinal_ThTo_SimpEffect_NoSound["Congruency_C:Validity_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Acc_ModFinal_ThTo_SimpEffect_NoSound["Congruency_C:Validity_C", "Pr...z.."])`).



## Discussion 

In this study, as in the previous one, we tested the hypothesis that threat is likely to enhance goal-driven attentional priorities. 
To do so, we employed a spatial cueing task in which distress screams were unpredictably heard, aiming to induce a state of threat in the participants. 
Marginal results suggest an effect of threat on response accuracy, but not on response times. 
Specifically, when screams were heard during a block of trials, participants tended to place more weight on the congruency of the cue in their responses. 
Thus, when the distractor cue is congruent with the attentional set, cue validity has a greater influence on task errors. In contrast, when the distractor cue is incongruent with the attentional set, the validity effect appears reduced under threat compared to the control condition. 
The attentional capture hypothesis posits that an irrelevant distractor is only captured when its attributes match those defined as priorities in the attentional set. 
The results of this experiment suggest that threat amplifies this phenomenon of attentional capture, at least in terms of response accuracy, with a simple search task. 
Thus, this experiment provides further evidence for the Arousal-Biased Competition model [@matherArousalBiasedCompetitionPerception2011], which proposes that anxious arousal strengthens goal-driven attentional priorities. 


Furthermore, the exploratory analyses from both studies provide additional support for the Arousal Biased Competition theory. 
Specifically, by aggregating data from the *Threat* and *Vocalize* blocks, our results show an effect of sound unpredictability on both response accuracy and response times. 
Although some effects are marginal, the impact of sounds on attentional capture remains consistent with our pre-registered hypotheses. 
Thus, the unpredictable occurrence of sounds during the task enhances the effect of attentional capture, such that response speed and accuracy are more influenced by a congruent cue for task goals, and less influenced by an incongruent cue.


The results of these two experiments suggest enhanced attentional selectivity when combining data from the *Threat* and *Vocalise* blocks. 
However, it is surprising that the effect observed in @normandDoesEvaluativePressure2014 was not replicated when comparing only the *Control* and *Threat* blocks. 
This non-replication may stem from differences in how threat was conceptualized. 
In our experiments, using the Threat of Scream paradigm, the threat consisted of unpleasant sounds occurring randomly throughout the task. 
In this way, the uncertainty of the screams contributed to increased self-reported anxiety, without giving participants the ability to reduce the threat.
In contrast, Normand et al. (2014) framed the threat as directly related to task performance, leading participants to believe they could somewhat influence the threat. 
Specifically, participants in the self-evaluative threat condition were told that the spatial cueing task was strongly linked to intellectual ability, creating a perception of contingency between task performance and intelligence.
While this procedure threatened participants based on their performance, it also provided them with control over the threat, because this led them believed that they could limit the threat by performing well on the spatial cueing task.


Research suggests that the degree of control individuals have over a threat influences their responses to it [@herbertEffectVaryingDanger2020]. 
Specifically, having control over an aversive event reduces its impact intensity [@crombezItBetterHave2008; @salomonsPerceivedControllabilityModulates2004; @thompsonWillItHurt1981], as well as anticipatory stress, measured through physiological or self-reported assessments [for a review, see @millerControllabilityHumanStress1979]. 
Additionally, studies indicate that attentional and behavioral responses differ depending on the controllability of a threat [@georgiadesRelationshipWorryAttentional2021; @herbertEffectVaryingDanger2020; @notebaertTraitAnxietyAlignment2020]. 
While the procedure of Normand et al. (2014) offers control over the threat, which participants in our Threat of Scream condition do not have, it is conceivable that the non-replication of the original study is primarily due to the uncontrollability of the threat in our studies. 
To test the influence of threat controllability on attentional priorities, a new study was conducted.












# Supplementary

## Supplementary : Study 2 {#Supplementary-Study-2}

### Results 

#### Effectiveness of anxiety induction

As in Study 1, self-reported anxiety at the end of the *Threat* block was significantly higher than at the beginning of the experiment (*b* =
`r round(Report_Evol_lm01_Location$coefficients[1,1], digits = 2)`,
*F*(`r Report_Evol_lm01_Location$df[1]`, `r Report_Evol_lm01_Location$df[2]`) =
`r round(Report_Evol_lm01_Location$coefficients[1,3], digits = 2)`, *p* `r pValue(Report_Evol_lm01_Location$coefficients[1,4])`). Significant increases in initial anxiety levels were also observed in the *Control* and *Vocalize* blocks (respectively *b* =
`r round(Report_Evol_lm02_Location$coefficients[1,1], digits = 2)`,
*F*(`r Report_Evol_lm02_Location$df[1]`, `r Report_Evol_lm02_Location$df[2]`) =
`r round(Report_Evol_lm02_Location$coefficients[1,3], digits = 2)`, *p* `r pValue(Report_Evol_lm02_Location$coefficients[1,4])` and *b* = `r round(Report_Evol_lm03_Location$coefficients[1,1], digits = 2)`,
*F*(`r Report_Evol_lm03_Location$df[1]`, `r Report_Evol_lm03_Location$df[2]`) =
`r round(Report_Evol_lm03_Location$coefficients[1,3], digits = 2)`, *p* `r pValue(Report_Evol_lm03_Location$coefficients[1,4])`). 

The increase in anxiety was significantly higher in the *Threat* block (*M* =
`r round(Evolution_table_Location[[2,"Mean"]], digit = 2)`, *ET* =
`r round(Evolution_table_Location[[2,"SD"]], digit = 2)`) than in the *Control* block (*M* =
`r round(Evolution_table_Location[[1,"Mean"]], digit = 2)`, *ET* =
`r round(Evolution_table_Location[[1,"SD"]], digit = 2)`, *b* = `r round(coefs_Evol_lmer01_Location["BlockThreat", "Estimate"], digits = 2)`,
*t*(`r round(get_df(Evol_lmer01_Location, type = "ml1")[["BlockThreat"]], digit =0)`) =
`r round(coefs_Evol_lmer01_Location["BlockThreat", "t.value"], digits = 2)`, *p* 
`r pValue(coefs_Evol_lmer01_Location["BlockThreat", "Pr...t.."])`), or the *Vocalize* one (*M* =
`r round(Evolution_table_Location[[3,"Mean"]], digit = 2)`, *ET* =
`r round(Evolution_table_Location[[3,"SD"]], digit = 2)`, *b* = `r round(coefs_Evol_lmer02_Location["BlockToon", "Estimate"], digits = 2)`,
*t*(`r round(get_df(Evol_lmer02_Location, type = "ml1")[["BlockToon"]], digit =0)`) =
`r round(coefs_Evol_lmer02_Location["BlockToon", "t.value"], digits = 2)`, *p* `r pValue(coefs_Evol_lmer02_Location["BlockToon", "Pr...t.."])`). Regarding anxiety increases, no significant difference was found between the *Control* and *Vocalize* blocks (*b* = `r round(coefs_Evol_lmer01_Location["BlockToon", "Estimate"], digits = 2)`,
*t*(`r round(get_df(Evol_lmer01_Location, type = "ml1")[["BlockToon"]], digit =0)`) =
`r round(coefs_Evol_lmer01_Location["BlockToon", "t.value"], digits = 2)`, *p* `r pValue(coefs_Evol_lmer01_Location["BlockToon", "Pr...t.."])`).

This time again, participants rated screams as significantly more threatening (*M* =
`r round(mean(df_Survey_Wide_Location$Threat_Scream_Mean), digit = 2)`, *ET* =
`r round(sd(df_Survey_Wide_Location$Threat_Scream_Mean), digit = 2)`) compared to vocalizations (*M* =
`r round(mean(df_Survey_Wide_Location$Threat_Vocal_Mean), digit = 2)`, *ET* =
`r round(sd(df_Survey_Wide_Location$Threat_Vocal_Mean), digit = 2)`, *b* =
`r round(Report_Threatening_Sounds_Location$coefficients[1,1], digits = 2)`,
*F*(`r Report_Threatening_Sounds_Location$df[1]`, `r Report_Threatening_Sounds_Location$df[2]`) =
`r round(Report_Threatening_Sounds_Location$coefficients[1,3], digits = 2)`, *p* `r pValue(Report_Threatening_Sounds_Location$coefficients[1,4])`).
Moreover, the threatening nature of screams significantly predicted the self-reported anxiety increase following the *Threat* block (*b* = `r round(Report_Anxiety_Screams_lmMean_Location$coefficients[2,1], digits = 2)`,
*F*(`r Report_Anxiety_Screams_lmMean_Location$df[1]-Report_Anxiety_Screams_lmMean_Location[["fstatistic"]][["numdf"]]`, `r Report_Anxiety_Screams_lmMean_Location$df[2]`) =
`r round(Report_Anxiety_Screams_lmMean_Location$coefficients[2,3], digits = 2)`, *p* `r pValue(Report_Anxiety_Screams_lmMean_Location$coefficients[2,4])`, η² = `r round(rsq(Anxiety_Screams_lmMean_Location), digit = 3)`), just as the threatening nature of vocalizations significantly predicted the self-reported anxiety increase in the *Vocalize* block (*b* = `r round(Report_Anxiety_Vocaliz_lmMean_Location$coefficients[2,1], digits = 2)`,
*F*(`r Report_Anxiety_Vocaliz_lmMean_Location$df[1]-Report_Anxiety_Vocaliz_lmMean_Location[["fstatistic"]][["numdf"]]`, `r Report_Anxiety_Vocaliz_lmMean_Location$df[2]`) =
`r round(Report_Anxiety_Vocaliz_lmMean_Location$coefficients[2,3], digits = 2)`, *p* `r pValue(Report_Anxiety_Vocaliz_lmMean_Location$coefficients[2,4])`, η² = `r round(rsq(Anxiety_Vocaliz_lmMean_Location), digit = 3)`).


#### Contingent capture effect

On response times, our analyses reveal a main effect of cue validity, consistent with Study 1 (*b* = `r round(coefs_RT_ModFinal_02["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Validity_C", "t.value"], digits = 2)`, *p* `r pValue(coefs_RT_ModFinal_02["Validity_C", "Pr...t.."])`), as well as a main effect of congruency (*b* = `r round(coefs_RT_ModFinal_02["Congruency_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Congruency_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Congruency_C", "t.value"], digits = 2)`, *p* `r pValue(coefs_RT_ModFinal_02["Congruency_C", "Pr...t.."])`). 
Additionally, we find an interaction between cue *Congruency* and *Validity* on response times (*b* = `r round(coefs_RT_ModFinal_02["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_ModFinal_02, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`) =
`r round(coefs_RT_ModFinal_02["Congruency_C:Validity_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_ModFinal_02["Congruency_C:Validity_C", "Pr...t.."])`). 
This interaction, in line with the contingent capture effect, indicates that the cueing effect is stronger in congruent trials (*b* = `r round(coefs_RT_SimpEffect_Cong["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_SimpEffect_Cong, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_SimpEffect_Cong["Validity_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_SimpEffect_Cong["Validity_C", "Pr...t.."])`) than in incongruent trials (*b* = `r round(coefs_RT_SimpEffect_Incong["Validity_C", "Estimate"], digits = 2)`,
*t*(`r round(get_df(RT_SimpEffect_Incong, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_RT_SimpEffect_Incong["Validity_C", "t.value"], digits = 2)`, *p* `r  pValue(coefs_RT_SimpEffect_Incong["Validity_C", "Pr...t.."])`).


```{r FigureContingentCapture-Locate, fig.cap = "(ref:Title-FigureContingentCapture-Locate)", fig.align = 'center', echo=FALSE, warning=FALSE}

# APA Style Line Graph with a Single Legend for Congruency
ggplot(Mean_RT_Contingent_Capture, aes(x = Cueing_Validity, y = mean, color = Congruency, linetype = Congruency, group = Congruency)) +
  
  # Points and Error Bars
  geom_point(position = position_dodge(width = 0.2), size = 2) +  # Larger point size for clarity
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.1, position = position_dodge(width = 0.2), size = 0.8) +  # Error bars with proper width
  
  # Line
  geom_line(position = position_dodge(width = 0.2), size = 1) +  # Slightly thicker lines for APA clarity
  
  # Labels for X and Y axis
  xlab("Cueing Validity") +
  ylab("Response Time (in milliseconds)") +
  
  # Custom X-axis labels
  scale_x_discrete(limits = c("Valid", "Invalid"), 
                   labels = c("Valid", "Invalid")) +
  
  # Custom color and linetype scale, with combined legend
  scale_color_manual(
    values = c("green4", "red", "black"),  # Custom colors for the legend
    labels = c("Congruent", "Incongruent", "No Cue")  # Custom labels
  ) +
  
  # Custom linetype scale (linked to same variable)
  scale_linetype_manual(
    values = c("solid", "solid", "dashed"),  # Linetypes
    labels = c("Congruent", "Incongruent", "No Cue")  # Ensure labels match color legend
  ) +
  
  # Combine the color and linetype legends into one
  guides(
    color = guide_legend(title = "Cueing Congruency", 
                         override.aes = list(linetype = c("solid", "solid", "dashed"))),
    linetype = "none"  # Hide the separate linetype legend
  ) +
  
  # APA style theme adjustments
  theme(legend.position = "top",             # Remove legend 
        panel.grid.major = element_blank(),   # Remove grid lines
        panel.grid.minor = element_blank(),   # Remove minor grid lines
        axis.line = element_line(size = 0.5, color = "black") # Add axis lines for clarity
  )

```

(ref:Title-FigureContingentCapture-Locate) **Study 2: Involuntary contingent capture of attention**.  


On response accuracy, our analyses reveal a main effect of cue validity (*b* = `r round(coefs_Accuracy_ModFinal["Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Validity_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Validity_C", "Pr...z.."])`) as well as a main effect of cue congruency (*b* = `r round(coefs_Accuracy_ModFinal["Congruency_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Congruency_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Congruency_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Congruency_C", "Pr...z.."])`). 
Furthermore, the results show a significant interaction between *Validity* and *Congruency* (*b* = `r round(coefs_Accuracy_ModFinal["Congruency_C:Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_ModFinal, type = "ml1")[["Congruency_C:Validity_C"]], digit =0)`) =
`r round(coefs_Accuracy_ModFinal["Congruency_C:Validity_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_ModFinal["Congruency_C:Validity_C", "Pr...z.."])`) : the effect of validity on accuracy is greater in congruent trials (*b* = `r round(coefs_Accuracy_SimpEffect_Cong["Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_SimpEffect_Cong, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_Accuracy_SimpEffect_Cong["Validity_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_SimpEffect_Cong["Validity_C", "Pr...z.."])`) than in incongruent trials (*b* = `r round(coefs_Accuracy_SimpEffect_Incong["Validity_C", "Estimate"], digits = 2)`,
*Z*(`r round(get_df(Accuracy_SimpEffect_Incong, type = "ml1")[["Validity_C"]], digit =0)`) =
`r round(coefs_Accuracy_SimpEffect_Incong["Validity_C", "z.value"], digits = 2)`, *p* `r  pValue(coefs_Accuracy_SimpEffect_Incong["Validity_C", "Pr...z.."])`).



#### General Model Parameters for Response Time

The variance in response times is explained by the final model at `r round((perf_RT_ModFinal_02[1, "R2_conditional"]*100), digits = 1)`% par le modèle final. 
Furthermore, random effects from the `Participant` factor alone account for `r round((perf_RT_ModFinal_02[1, "ICC"]*100), digits = 1)`% of the variability in the data, while fixed effects account for only `r round((perf_RT_ModFinal_02[1, "R2_marginal"]*100), digits = 1)`% of the variance. 
In these analyses, the thresholds for data outlier removal correspond to lever values greater than 0.071 and Cook's distances greater than 0.050. 
The final dataset consists of *k* = `r nrow(df_Locate_NoNeut_Prereg_Excl_RT_Out)` trials.


#### General Model Parameters for Accuracy

For response accuracy, `r round((perf_Accuracy_ModFinal[1, "R2_conditional"]*100), digits = 1)`% of the variance in the data is explained by the model. Random effects and fixed effects account for `r round((perf_Accuracy_ModFinal[1, "ICC"]*100), digits = 1)`% and `r round((perf_Accuracy_ModFinal[1, "R2_marginal"]*100), digits = 1)`% of the variability, respectively. 
In these analyses, the thresholds for data outlier removal correspond to lever values greater than 0.80 and Cook's distances greater than 1.00. The final dataset consists of *k* = `r nrow(df_Locate_NoNeut_Prereg_Excl_Acc_Out)` trials.


#### Exploratory analyses: General Model Parameters for Response Times

To analyze the influence of sound occurrence on response times, the final dataset included *k* = `r nrow(df_Locate_ThTo_Prereg_Excl_Out)` trials. 
Our analyses reveal that `r round((perf_RT_ModFinal_ThTo_02[1, "R2_conditional"]*100), digits = 1)`% of the total variance in the data is explained by our model, which includes both fixed and random effects. Furthermore, `r round((perf_RT_ModFinal_ThTo_02[1, "ICC"]*100), digits = 1)`% of the variance is explained by the random effects, while `r round((perf_RT_ModFinal_ThTo_02[1, "R2_marginal"]*100), digits = 1)`% is explained by the fixed effects.
In these analyses, the thresholds for data outlier removal correspond to lever values greater than 0.045 and Cook's distances greater than 0.028.


#### Exploratory analyses: General Model Parameters for Acccuracy

The final dataset used to analyze the influence of sound occurrence on response accuracy consists of *k* = `r nrow(df_Locate_ThTo_Prereg_Excl_Acc_Out)` trials. 
Our analyses reveal that `r round((perf_Accuracy_ModFinal_ThTo[1, "R2_conditional"]*100), digits = 1)`% of the total variance in the data is explained by our model, which includes both fixed and random effects. Furthermore, `r round((perf_Accuracy_ModFinal_ThTo[1, "ICC"]*100), digits = 1)`% of the variance is explained by the random effects, while `r round((perf_Accuracy_ModFinal_ThTo[1, "R2_marginal"]*100), digits = 1)`% is explained by the fixed effects.
In these analyses, the thresholds for data outlier removal correspond to lever values greater than 0.66 and Cook's distances greater than 1.40.
